Time delay estimation in unresolved lensed quasars

### 2020-12-28 paper/aa-package/paper.tex
Update paper.tex

  \abstract
  % context heading (optional)
  % {} leave it empty if necessary  
   {
   	Early universe (EU) measurements and late universe (LU) observations have resulted in a tension on the estimated value of the Hubble parameter $H_0$. Time-delay cosmography offers an alternative method to measure $H_0$. In this respect, the H0LiCoW collaboration has reported a $2.4\%$ measurement of $H_0$ compatible with LU observations, increasing the tension at the $5.3 \sigma$ level. Whereas, TDCOSMO+SLACS has reported a $5\%$ measurement of $H_0$ in agreement with both EU and LU estimates, showing the need to collect more data in order to reduce the error in the $H_0$ estimation.
   	}
  % aims heading (mandatory)
   { 
   	In time-delay cosmography, the fractional error on $H_0$ is directly related to the error of relative time delays measurements and it linearly decreases with the number of lensed systems considered. Therefore, in order to reduce it, more lensed systems should be analysed and, possibly, with a regular and long-term monitoring, of the order of years. This cannot be achieved with big telescopes, due to the huge amount of observational requests they have to fulfill. On the other hand, small/medium-sized telescopes are present in a much larger number and are often characterized by more versatile observational programs. However, the limited resolution capabilities of such instruments and their often not privileged geographical location prevent them from providing well-separated images of the same lensed source. 
   	\\
   	In this work, we present a novel approach to estimate the time-delay in unresolved lensed quasar systems. Our proposal is further motivated by recent developments in discovering more unresolved strongly-lensed QSO systems.
	}
  % methods heading (mandatory)
   {Our method uses ...}
  % results heading (mandatory)
   {...}
  % conclusions heading (optional), leave it empty if necessary 
   {}

@@

The Hubble parameter $H_0$ quantifies the current expansion rate of the Universe. The measured value of $H_0$ from different observations have led to a tension. In particular, EU observations [\cite{planck}] have measured $H_0 = 67.4 \pm 0.5$ km s$^{-1}$ Mpc$^{-1}$, whereas, LU observations [\cite{cepheid_H0}] give $H_0 = 74.03 \pm 1.42$ km s$^{-1}$ Mpc$^{-1}$, resulting in a tension of about $4.4\sigma$.

@@

As first pointed out in [\cite{Refsdal1964}], time-delay cosmography offers an alternative way of determining the Hubble parameter: the light rays coming from a distant source, e.g. a quasar, can be deflected by the gravitational field of an intervening massive object, e.g. a galaxy. If the field is strong enough, multiple images of the same source are observed and, by tracking the light intensity of each image over time, a \emph{light curve} is obtained. Light curves associated with different images will exhibit a mutual time-delay ($\Delta T$), due to the different paths that the photons have travelled. As shown in [\cite{Refsdal1964}], this time-delay is related to the Hubble parameter as $H_0 \propto 1/\Delta T$. The major results obtained via time-delay cosmography come from the H0LiCOW collaboration [\cite{H0licow_XIII}], which has found $H_0 = 73.3_{-1.8}^{+1.7}$ km s$^{-1}$ Mpc$^{-1}$ from a sample of six lensed quasars monitored by the COSMOGRAIL  project [\cite{Cosmograil2020}], enhancing the tension up to $5.3 \sigma$. However, a more recent analysis of 40 strong gravitational lenses, from TDCOSMO+SLACS [\cite{tdcosmo}], has found $H_0 = 67.4_{-3.2}^{+4.1}$ km s$^{-1}$ Mpc$^{-1}$, relaxing the Hubble tension and demonstrating the importance of understanding the lenses mass density profiles. This scenario motivates further studies aimed at improving the precision in the $H_0$ estimation. 

@@

To date, a sample of about $220$ GLQs is available\footnote{https://research.ast.cam.ac.uk/lensedquasars/index.html}. However, only a very small subset, with well separated multiple images, has been used to measure $H_0$. The reason relies on the fact that it is easier and safer to extract information from such systems, and consequently reduce the error on $H_0$.

@@

In this respect, Fig. \ref{fig:qsosepmagboth} (bottom) shows the magnitude of the multiple images versus the maximum image separation for the known GLQs. Systems being part of the grey region, which represent the 70$\%$ of the total sample (as shown in the up-right histogram), have a maximum image separation below 2 arcsec. 

@@

In addition, [\cite{shu2020discovering}] have presented a new method to find GLQs from unresolved light curves. Such systems have therefore even smaller separation between multiple images. 

@@

This would make big telescopes the ideal instruments to perform lensed quasars monitoring, both in light of their high angular resolution and the geographical areas they are placed in, where the effects of atmospheric turbulence are less prominent. However, because of the time scales of the intrinsic variations of the source, such observation campaigns should last years [\cite{Cosmograil2020}]. Therefore, due to the amount of observational requests that big telescopes have to fulfil, they can not be employed for these purposes. On the other hand, small/medium sized telescopes ($1$-$2$m) [\cite{Borgeest1996}] can be used. Unfortunately, their already reduced angular resolution is further worsened by their less privileged geographical positions, in terms of atmospheric seeing, which can reach 3 arcsec [\cite{karttunen2016fundamental}]. 

>>>>

  \abstract
  % context heading (optional)
  % {} leave it empty if necessary  
   {
   	Early universe (EU) measurements and late universe (LU) observations have resulted in a tension on the estimated value of the Hubble parameter $H_0$. Time-delay cosmography offers an alternative method to measure $H_0$. In this respect, the H0LiCoW collaboration has reported a $2.4\%$ measurement of $H_0$ compatible with LU observations, increasing the tension at the $5.3 \sigma$ level. Whereas, TDCOSMO+SLACS has reported a $5\%$ measurement of $H_0$ in agreement with both EU and LU estimates, showing the need to collect more data in order to reduce the uncertainty in the $H_0$ estimation.
   	}
  % aims heading (mandatory)
   { 
   	In time-delay cosmography, the fractional uncertainty on $H_0$ is directly related to the uncertainty on relative time delays measurements and it linearly decreases with the number of lensed systems considered. Therefore, in order to reduce it, more lensed systems should be analysed and, possibly, with a regular and long-term monitoring, of the order of years. This cannot be achieved with big telescopes, due to the huge amount of observational requests they have to fulfill. On the other hand, small/medium-sized telescopes are present in a much larger number and are often characterized by more versatile observational programs. However, the limited resolution capabilities of such instruments and their often not privileged geographical location may prevent them from providing well-separated images of the same lensed source. 
   	\\
   	In this work, we present a novel approach to estimate the time-delay in unresolved lensed quasar systems. Our proposal is further motivated by recent developments in discovering more unresolved strongly-lensed QSO systems.
	}
  % methods heading (mandatory)
   {Our method uses ...}
  % results heading (mandatory)
   {...}
  % conclusions heading (optional), leave it empty if necessary 
   {}

@@

The Hubble parameter $H_0$ quantifies the current expansion rate of the Universe. The measured values of $H_0$ from different observations led to a tension. In particular, EU observations [\cite{planck}] have measured $H_0 = 67.4 \pm 0.5$ km s$^{-1}$ Mpc$^{-1}$, whereas, LU observations [\cite{cepheid_H0}] give $H_0 = 74.03 \pm 1.42$ km s$^{-1}$ Mpc$^{-1}$, resulting in a tension of about $4.4\sigma$.

@@

As first pointed out in [\cite{Refsdal1964}], time-delay cosmography offers an alternative way of determining the Hubble parameter: the light rays coming from a distant source, e.g. a quasar, can be deflected by the gravitational field of an intervening massive object, e.g. a galaxy. If the field is strong enough, multiple images of the same source are observed and, by tracking the light intensity of each image over time, a \emph{light curve} is obtained. Light curves associated with different images will exhibit a mutual time-delay ($\Delta T$), due to the different paths that the photons have travelled. As shown in [\cite{Refsdal1964}], this time-delay is related to the Hubble parameter as $H_0 \propto 1/\Delta T$. The major results obtained via time-delay cosmography come from the H0LiCOW collaboration [\cite{H0licow_XIII}], who found $H_0 = 73.3_{-1.8}^{+1.7}$ km s$^{-1}$ Mpc$^{-1}$ from a sample of six lensed quasars monitored by the COSMOGRAIL  project [\cite{Cosmograil2020}], enhancing the tension up to $5.3 \sigma$. However, a more recent analysis of 40 strong gravitational lenses, from TDCOSMO+SLACS [\cite{tdcosmo}], found $H_0 = 67.4_{-3.2}^{+4.1}$ km s$^{-1}$ Mpc$^{-1}$, relaxing the Hubble tension and demonstrating the importance of understanding the lenses mass density profiles. This scenario motivates further studies aimed at improving the precision in the $H_0$ estimation. 

@@

To date, a sample of about $220$ GLQs is available\footnote{https://research.ast.cam.ac.uk/lensedquasars/index.html}. However, only a very small subset, with well separated multiple images, has been used to measure $H_0$. The reason is that it is easier and safer to extract information from such systems, and consequently reduce the error on $H_0$.

@@

In this respect, Fig. \ref{fig:qsosepmagboth} (bottom) shows the magnitude of the multiple images versus the maximum image separation for the known GLQs. Systems being part of the grey region, which represent 70$\%$ of the total sample (as shown in the up-right histogram), have a maximum image separation below 2 arcsec. 

@@

In addition, [\cite{shu2020discovering}] presented a new method to find GLQs from unresolved light curves. Such systems have therefore even smaller separation between multiple images. 

@@

This would make big telescopes the ideal instruments to perform lensed quasars monitoring, both in light of their high angular resolution and the geographical areas they are placed in, where the effects of atmospheric turbulence are less prominent. However, because of the time scales of the intrinsic variations of the sources, such observation campaigns should last years [\cite{Cosmograil2020}]. Therefore, due to the amount of observational requests that big telescopes have to fulfil, they hardly can be employed for these purposes. On the other hand, small/medium sized telescopes ($1$-$2$m) [\cite{Borgeest1996}] can be used. Unfortunately, their already reduced angular resolution is further worsened by their often less privileged geographical positions, in terms of atmospheric seeing, which can reach 3 arcsec [\cite{karttunen2016fundamental}]. 
